### 2023-02-01 article3.qmd
update

@fig-uvp-vertical-profiles shows that all floats measure similar patterns in the vertical profiles of particles concentrations in all size classes. In the 102-512 $\mu$m range, all 4 floats measured a local maximum between 500 and 1500 m, likely due to both sinking and remineralization processes. Above 512 $\mu$m, the relative abundance is maximum along a depth track centered on August. It can be seen among all size classes that the water column is being drained proportionally to the class size.

@@

@@

## plot absolute concentrations
up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
  #group_by(size) |>
  ggplot() + facet_grid(wmo~size) + theme_bw() +
  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
  #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
  #scale_colour_viridis(option = 'H') +
  scale_y_reverse() +
  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
  #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
  labs(colour = '#/L', x = 'Month', y = 'Depth (m)')
  #theme(legend.title = element_text(angle = -90))
  #guides(fill = guide_legend(title.theme = element_text(size = 200)))
  #+ theme(legend.position = 'none')
  #geom_path(data = particles_bis, aes(juld, MLD), colour = "black") +
  #theme(legend.position = 'none') #+
  #geom_hline(yintercept = c(200,500,1000), linetype = 'dotted') #+
  #guides(colour = guide_legend(title.position = "bottom"))

@@

### Where is the bloom?

```{r, sat data}

#chla_sat <- vroom::vroom('DATA/SATELLITE/chla_sat.csv')
chla_sat <- arrow::read_feather('DATA/SATELLITE/chla_sat.feather') |> mutate(DOY = yday(time))
sat_data_bbox <- filter(chla_sat, lat >= 55.5, lat <= 60, lon >= -55, lon <= -45) |> group_by(DOY) |> drop_na(CHL)

# sat data
test <- sat_data_bbox |> group_by(DOY) |> summarize(npixels = n(), nabove5 = length(which(CHL > 5)), nabove2 = length(which(CHL > 2)), r5 = nabove5/npixels)
ggplot(test, aes(DOY, r5)) + geom_point() + geom_path()

# argo data
npq_data <- vroom('DATA/ARGO/npq_corrected_data.csv')  
npq_filtered <- npq_data |> filter(pres > 0, pres <= 20, juld < '2022-12-31') |> mutate(DOY = yday(juld)) |> drop_na(chla)

npq_argo <- npq_filtered |> group_by(wmo, cycle) |> summarize(DOY = unique(DOY), mean_chla = mean(chla_npq, na.rm=T), median_chla = median(chla_npq, na.rm=T))

ggplot(npq_argo, aes(x = DOY, y = mean_chla, colour = factor(wmo))) + geom_point() + geom_path() + facet_wrap(~wmo)

@@

The use of @eq-guidi-article3 to derive carbon fluxes from PSD was firstly implemented by @Guidi2008-vp who used a minimization procedure between sediment traps data and estimated fluxes using PSD from several previous versions of the UVP to find the optimal value of $A$ and $b$ (i.e. 12.5 and 3.81). @Iversen2010-an and @Fender2019-ou followed the same procedure and concluded that the couple of parameters found by [@Guidi2008-vp] could not be used globally at the expense of greatly over- or underestimating POC fluxes, depending on the sampling region. In addition, the formulation of such a model implies that $A$ and $b$ would be valid at global scale, at all depths, for all UVP models and across all size classes which is arguable [@Bisson2022-bt]. @eq-guidi-article3 also assumes that the flux is only size-dependent though the main driver for the settling of aggregates is still debated [@Iversen2020-kb; @Iversen2022-qj]. Due to those reasons, we have chosen to firstly compare $F_{small}$ and $F_{big}$ with UVP6 particles concentrations. For this, we computed the particles density distribution maximum using a smoothing gaussian kernel (see R package ggvis) for each class size at each parking depth and at each cycle. Then, we computed the distance correlation between OST-derived carbon fluxes and UVP size classes maxima. The distance correlation \[REF\] was preferred to the well known Pearson's correlation due to uncertainties in the linear relationship of, respectively, small classes and $F_{big}$, and big classes and $F_{small}$.

```{r, essai sur ridgeplot}
library(ggridges)
tmp <- uvp_data |> filter(park_depth == '1000 m', wmo == 6904240, conc < 20)
#tmp <- uvp_data |> filter(park_depth == '1000 m', wmo == 6904240, size_class %in% c('NP_Size_102', 'NP_Size_128', 'NP_Size_161', 'NP_Size_203', 'NP_Size_256', 'NP_Size_323'))

# ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + geom_density_ridges(scale = 5) + stat_density_ridges(quantile_lines=T, quantiles = 2) + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
#   theme_bw() + scale_color_brewer(palette = 'Dark2')
# 
# # ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + geom_density_ridges() + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) 
# #   
# # ggplot(tmp, aes(y = factor(cycle), x = conc)) + geom_density_ridges_gradient() + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
# #    scale_fill_viridis_c(name = "Temp. [F]", option = "C")
# 
# ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + stat_density_ridges(quantile_lines=T, quantiles = 2) + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
#   theme_bw() 
# 
# ggplot(tmp, aes(y = factor(cycle), x = conc)) + stat_density_ridges(quantile_lines=T, quantiles = 2) + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
#   theme_bw() + facet_wrap(~size_class) + xlim(c(0, 10)) 
# 
# ggplot(tmp, aes(y = factor(cycle), x = conc)) + stat_density_ridges(quantile_lines=T, quantiles = 2, stat="identity") + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
#   theme_bw() + facet_wrap(~size_class, scales = "free")
# 
ggplot(tmp, aes(y = factor(cycle), x = conc, height = ..density..)) + geom_density_ridges(scale = 5, stat = "density", rel_min_height = 0.01) + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
  theme_bw() + facet_wrap(~size_class, scales = "free")
# 
# # test 
# tmp2 <- uvp_data |> dplyr::group_by(cycle, size_class, park_depth, wmo) |> 
#   group_modify(~ ggplot2:::compute_density(.x$conc, NULL)) |>
#   rename(conc = x)
# 
# ggplot(tmp2, aes(x = conc, y = cycle, height = density)) + 
#   geom_density_ridges(stat = "identity") + facet_wrap(~size_class, scales = "free")

# 
# ggplot(tmp, aes(y = factor(cycle), x = conc, height = ..density..)) + stat_density_ridges(scale = 5) + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
#   theme_bw() + facet_wrap(~size_class, scales = "free")
# 
# ggplot(tmp, aes(y = factor(cycle), x = conc, height = ..density..)) + stat_density_ridges(scale = 5, quantile_lines = T, quantiles = 2) + labs(x = 'abundance distribution (#/L)', y = 'cycle')  +
#   theme_bw() + facet_wrap(~size_class, scales = "free")
# 
# ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + geom_ridgeline() + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
#   theme_bw() + scale_color_brewer(palette = 'Dark2')

# ggplot(tmp, aes(x = factor(cycle), y = conc, fill = size_class)) + geom_boxplot() + labs(y = 'abundance distribution (#/L)', x = 'cycle')  +
#   theme_bw() + scale_color_brewer(palette = 'Dark2') + ylim(c(0,15)) + geom_jitter()
# 
# ggplot(tmp, aes(x = factor(cycle), y = conc, fill = size_class)) + geom_violin() + labs(y = 'abundance distribution (#/L)', x = 'cycle')  +
#   theme_bw() + scale_color_brewer(palette = 'Dark2') + ylim(c(0,15))

# ggplot(tmp, aes(y = factor(cycle), x = conc, fill = size_class)) + geom_density_ridges(alpha=1, stat="binline", bins=50)  + labs(x = 'abundance distribution (#/L)', y = 'cycle') + xlim(c(0,10)) +
#   theme_bw() + scale_color_brewer(palette = 'Dark2')

@@

```{r, check with particles abundance}

particles <- all_floats |> filter(PhaseName == 'NPAR') |> select(juld, cycle, depth = pres, wmo, all_of(lpm_class)) |> drop_na(NP_Size_102) |> mutate(juld = as_date(juld))

particles_bis <- particles |> pivot_longer(cols = all_of(lpm_class), names_to = 'size_class', values_to = 'concentration')
particles_bis$size_class <- factor(particles_bis$size_class, levels = lpm_class)

ggplot(particles_bis |> filter(wmo == '1902578'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance', title = 'wmo 1902578') +
  facet_wrap(~size_class) + theme_bw()

ggplot(particles_bis, aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance') +
  facet_wrap(~size_class) + theme_bw()

ggplot(particles_bis |> filter(wmo == '6904240'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance', title = 'wmo 6904240') +
  facet_wrap(~size_class) + theme_bw()

# https://stackoverflow.com/questions/48424682/how-do-i-limit-the-range-of-the-viridis-colour-scale
ggplot(particles_bis |> filter(wmo == '6904240'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_colour_viridis(option = 'H', limits = c(-1,1), oob = scales::squish, na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance', title = 'wmo 6904240') +
  facet_wrap(~size_class) + theme_bw()

ggplot(particles_bis |> filter(wmo == '6904241'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance', title = 'wmo 6904241') +
  facet_wrap(~size_class) + theme_bw()

ggplot(particles_bis |> filter(wmo == '4903634'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance', title = 'wmo 4903634') +
  facet_wrap(~size_class) + theme_bw()

ggplot(particles_bis |> filter(wmo == '1902578'), aes(x = juld, y = depth, colour = log10(concentration))) + geom_point() + scale_color_viridis_c(option = 'turbo', na.value = 'transparent') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'abundance', title = 'wmo 1902578') +
  facet_wrap(~size_class) + theme_bw()

# # sum 3 first class
# particles_tmp <- particles |> dplyr::group_by(wmo, cycle, depth) |> summarize(juld = unique(juld), sum_NP = sum(NP_Size_102, NP_Size_128, NP_Size_161)) |> filter(depth > 100)
# 
# ggplot(particles_tmp, aes(x = juld, y = depth, colour = log(sum_NP))) + geom_point() + scale_color_viridis(option = 'H') +
#   scale_y_reverse(limits = c(2000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))

# ggplot(particles, aes(x = juld, y = depth, colour = log(NP_Size_102))) + geom_point() + scale_color_viridis() +
#   scale_y_reverse(limits = c(2000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))

# ggplot(vertical_profiles, aes(x = juld, y = depth, colour = pred_flux_without_depth)) + geom_point() + scale_color_viridis() +
#   scale_y_reverse(limits = c(2000,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = TeX('predicted $F_{small}~(mg~m^{-2}~d^{-1})$'))

@@

```{r, check drift per class}

tmp <- drifting_profiles |> pivot_longer(cols = all_of(lpm_class[2:16]), names_to = "size_class")
tmp$park_depth <- factor(tmp$park_depth, levels = c('200 m', '500 m', '1000 m'))
tmp$size_class <- factor(tmp$size_class, levels = lpm_class)

tmp2 <- tmp |> group_by(park_depth, wmo, cycle, size_class) |> summarize(median_conc = median(value), min_juld = min(juld))

ggplot(tmp2 |> filter(wmo == 6904241, park_depth != '200 m'), aes(x = min_juld, y = median_conc, colour = factor(park_depth))) + geom_point() + geom_path() + facet_wrap(~size_class, scales = 'free') + theme_bw() + labs(title = 'WMO 6904241')

@@

```{r, check pulse export}

>>>>

@fig-uvp-vertical-profiles shows that all floats measure similar patterns in the vertical profiles of particles concentrations in all size classes. In the 102-512 $\mu$m range, all 4 floats measured a local maximum between 650 m and 1050 m. Above 512 $\mu$m, the relative abundance is maximum along a depth track centered on August. It can be seen among all size classes that the water column is being drained proportionally to the class size.

@@

# stats on the local maximum
# test <- particles_bis |> filter(depth > 500) |> group_by(wmo, cycle, size) |> # find depth of maximum relative abundance below 500 m between June and September
#   summarize(index_max = which.max(rel_conc), max_depth = depth[index_max], date = unique(juld)) |> 
#   filter(size < 512, date < '2022-10-01', date > '2022-05-31') |> # compute mean depth for each class size for each float
#   group_by(wmo, size) |> summarize(mean_depth = mean(max_depth))

@@

## plot absolute concentrations
up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
  #group_by(size) |>
  ggplot() + facet_grid(wmo~size) + theme_bw() +
  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
  #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
  #scale_colour_viridis(option = 'A', limits = c(-2,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
  #scale_colour_viridis(option = 'H') +
  scale_y_reverse() +
  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
  #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
  labs(colour = '#/L', x = 'Month', y = 'Depth (m)') #+
  #geom_hline(yintercept = c(600, 1100), linetype = 'dotted')
  #theme(legend.title = element_text(angle = -90))
  #guides(fill = guide_legend(title.theme = element_text(size = 200)))
  #+ theme(legend.position = 'none')
  #geom_path(data = particles_bis, aes(juld, MLD), colour = "black") +
  #theme(legend.position = 'none') #+
  #geom_hline(yintercept = c(200,500,1000), linetype = 'dotted') #+
  #guides(colour = guide_legend(title.position = "bottom"))

@@

### Chlorophyll data

```{r, chla data}

# argo data
npq_data <- vroom('DATA/ARGO/npq_corrected_data.csv')

# clean and smoothed chla data
d <- npq_data |> select(cycle, MLD, pres, juld, wmo, chla_npq) |> mutate(date = as_date(juld))
d <- d |> group_by(date, wmo, pres, cycle) |> summarize(mean_chla = mean(chla_npq, na.rm=T), mld = unique(MLD))
d <- d |> mutate(smoothed_chla = smooth(mean_chla))

d$wmo <- factor(d$wmo, levels = c(6904240,4903634,6904241,1902578))

# plot it
ggplot(d) + geom_point(aes(x = date, y = pres, colour = log10(smoothed_chla))) +
  scale_y_reverse(limits = c(200,0)) +
  #scale_colour_viridis(option = 'A') +
  scale_colour_cmocean(name = 'algae', oob = scales::squish, limits = c(-1,1), na.value = 'grey') +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'NPQ-corrected Chla') + facet_wrap(~wmo) + theme_bw()

# compute chla integration in the MLD and in the 0-100 m zone
d_chla <- d |> group_by(wmo, date) |> summarize(chla_0_100m = integrate(smoothed_chla, pres, from=0, to=100),
                                       mld = unique(mld),
                                       chla_0_mld = integrate(smoothed_chla, pres, from = 0, to=mld))

ggplot(d_chla) + geom_point(aes(date, chla_0_100m)) + facet_wrap(~wmo, scales = 'free') + theme_bw() + labs(x = 'Month', y = 'Integrated Chla 0-100 m (CHLA UNITS)')
#ggplot(d_chla) + geom_point(aes(date, chla_0_mld)) + facet_wrap(~wmo, scales = 'free')

```
### Mixed layer depth

```{r, mld data}
ggplot(d_chla) + geom_point(aes(date, mld)) + facet_wrap(~wmo, scales = 'free') + theme_bw() + labs(x = 'Month', y = 'Mixed layer depth (m)')
```

@@

The use of @eq-guidi-article3 to derive carbon fluxes from PSD was firstly implemented by @Guidi2008-vp who used a minimization procedure between sediment traps data and estimated fluxes using PSD from several previous versions of the UVP to find the optimal value of $A$ and $b$ (i.e. 12.5 and 3.81). @Iversen2010-an and @Fender2019-ou followed the same procedure and concluded that the couple of parameters found by [@Guidi2008-vp] could not be used globally at the expense of greatly over- or underestimating POC fluxes, depending on the sampling region. In addition, the formulation of such a model implies that $A$ and $b$ would be valid at global scale, at all depths, for all UVP models and across all size classes which is arguable [@Bisson2022-bt]. @eq-guidi-article3 also assumes that the flux is only size-dependent though the main driver for the settling of aggregates is still debated [@Iversen2020-kb; @Iversen2022-qj]. Due to those reasons, we have chosen to firstly compare $F_{small}$ and $F_{big}$ with UVP6 particles concentrations.

### Flux model(s)

```{r, flux model, include = FALSE}

flux_model <- function(boosting_rounds, which_flux){

    # select which flux to build the model
    if(which_flux == 'small'){
      cflux2 <- cflux |> mutate(flux = small_flux) 
    }else if(which_flux == 'big'){
      cflux2 <- cflux |> mutate(flux = big_flux)
    }else{
      cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
    }

  # training data
  tmp <- data_model |> select(flux, all_of(lpm_class))

  set.seed(101)
  # model
  model <- gbm(flux ~ ., data=tmp, distribution="gaussian",
  cv.folds=5,
  n.trees=boosting_rounds, shrinkage=0.005, interaction.depth=6,
  bag.fraction=0.5,
  n.cores=1
  )

  # get best model and thus best round number
  best <- gbm.perf(model)
  #print(best)

  # get features importance
  features_importance <- summary(model, n.trees=best, plotit = FALSE)

  return(list(features_importance = features_importance, model = model, best_tree = best))

# small_model <- flux_model(3000, 'small')
# big_model <- flux_model(3000, 'big')
# total_model <- flux_model(3000, 'total')
# 
# save(small_model, big_model, total_model, file = '/home/flo/dox/THESIS/DATA/MODEL/model_xgboost_data')

load(file = 'DATA/MODEL/model_xgboost_data')

@@

```{r, model predictions}

model_predict <- function(which_flux, mod, best_tree_number){

    # select which flux to build the model
    if(which_flux == 'small'){
      cflux2 <- cflux |> mutate(flux = small_flux) 
    }else if(which_flux == 'big'){
      cflux2 <- cflux |> mutate(flux = big_flux)
    }else{
      cflux2 <- cflux |> mutate(flux = small_flux + big_flux)
    }

  # predictions on mean spectra values
  uvp_spectra_average <- data_model |> dplyr::group_by(wmo, cycle, depth) |> dplyr::summarize(depth = unique(depth), true_flux = unique(flux),
                                                                                   NP_Size_102 = mean(NP_Size_102),
                                                                                   NP_Size_128 = mean(NP_Size_128),
                                                                                   NP_Size_161 = mean(NP_Size_161),
                                                                                   NP_Size_203 = mean(NP_Size_203),
                                                                                   NP_Size_256 = mean(NP_Size_256),
                                                                                   NP_Size_323 = mean(NP_Size_323),
                                                                                   NP_Size_406 = mean(NP_Size_406),
                                                                                   NP_Size_512 = mean(NP_Size_512),
                                                                                   NP_Size_645 = mean(NP_Size_645),
                                                                                   NP_Size_813 = mean(NP_Size_813),
                                                                                   NP_Size_1020 = mean(NP_Size_1020),
                                                                                   NP_Size_1290 = mean(NP_Size_1290),
                                                                                   NP_Size_1630 = mean(NP_Size_1630),
                                                                                   NP_Size_2050 = mean(NP_Size_2050)) |> ungroup()

  # prediction at 1000 m
  #spectra <- uvp_spectra_average |> select(true_flux, all_of(lpm_class[]))

  # prediction 
  pred_flux <- predict(mod, newdata = uvp_spectra_average, n.trees = best_tree_number)

  uvp_spectra_average$pred_flux <- pred_flux

  return(uvp_spectra_average)
}

# pred_small <- model_predict('small', small_model$model, small_model$best_tree) |> select(wmo, cycle, depth, true_small = true_flux, pred_small = pred_flux)
# pred_big <- model_predict('big', big_model$model, big_model$best_tree) |> select(true_big = true_flux, pred_big = pred_flux)
# pred_total <- model_predict('total', total_model$model, total_model$best_tree) |> select(true_total = true_flux, pred_total = pred_flux)
# 
# all_predictions <- cbind(pred_small, pred_big, pred_total) |> mutate(pred_total2 = pred_small + pred_big, anom = pred_total - pred_total2)
# 
# vroom::vroom_write(all_predictions, file = '/home/flo/dox/THESIS/DATA/MODEL/all_predictions.csv')
all_predictions <- vroom('DATA/MODEL/all_predictions.csv')

#### Features importance

```{r, features importance}

features <- small_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'small') |>
  bind_rows(big_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'big')) |>
  bind_rows(total_model$features_importance |> select(size = var, importance = rel.inf) |> mutate(model = 'total'))

rownames(features) <- NULL

features <- features |> mutate(size = size |> str_remove('NP_Size_') |> as.numeric())

ggplot(features) + geom_point(aes(x = as.factor(size), y = importance, colour = model), size = 5) + labs(x = 'Size', y = 'Features importance')

#### Predictions

```{r, plot predictions}

p_small <- ggplot(all_predictions) + geom_point(aes(true_small, pred_small)) + xlim(c(0,25)) + ylim(c(0,25)) + theme_bw() +
  annotate("text", x = 3, y = 23, label = paste0('R² = ' , round(R2_Score(all_predictions$true_small, all_predictions$pred_small), 2)))

p_big <- ggplot(all_predictions) + geom_point(aes(true_big, pred_big)) + xlim(c(0,50)) + ylim(c(0,50)) + theme_bw() +
  annotate("text", x = 5, y = 45, label = paste0('R² = ' , round(R2_Score(all_predictions$true_big, all_predictions$pred_big), 2)))

p_total <- ggplot(all_predictions) + geom_point(aes(true_total, pred_total)) + xlim(c(0,100)) + ylim(c(0,100)) + theme_bw() +
  annotate("text", x = 3, y = 80, label = paste0('R² = ' , round(R2_Score(all_predictions$true_total, all_predictions$pred_total), 2)))

(p_small + p_big) / p_total

# plot(test$true_flux, y=test$pred_flux, asp=1)
# print(R2_Score(test$pred_flux, test$true_flux))

#### Apply the total flux model on all vertical profiles

# make predictions
small_predicted_flux <- predict(small_model$model, newdata = vertical_profiles, n.trees = small_model$best_tree)
big_predicted_flux <- predict(big_model$model, newdata = vertical_profiles, n.trees = big_model$best_tree)
total_predicted_flux <- predict(total_model$model, newdata = vertical_profiles, n.trees = total_model$best_tree)

# add predictions to vertical data
vertical_profiles <- vertical_profiles |> mutate(small_pred = small_predicted_flux, big_pred = big_predicted_flux, total_pred = total_predicted_flux)

vertical_profiles$wmo <- factor(vertical_profiles$wmo, levels = c(6904240,4903634,6904241,1902578))

# pv_small <- ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(small_pred))) + geom_point() + scale_color_viridis(option = 'H') +
#   scale_y_reverse(limits = c(500,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = 'Predicted Fsmall') + facet_wrap(~wmo)
# 
# pv_big <- ggplot(vertical_profiles, aes(x = juld, y = depth, colour = log10(big_pred))) + geom_point() + scale_color_viridis(option = 'H') +
#   scale_y_reverse(limits = c(500,0)) +
#   scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
#   labs(x = 'Time', y = 'Predicted Fbig') + facet_wrap(~wmo)

pv_total <- ggplot(vertical_profiles, aes(x = juld, y = depth, colour = total_pred)) + geom_point() + #scale_color_viridis(option = 'H') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_colour_viridis(option = 'A', limits = c(quantile(all_predictions$true_total, probs = 0.25), quantile(all_predictions$true_total, probs = 0.75)), oob = scales::squish, na.value = 'grey', direction = -1) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'Predicted Ftotal') + facet_wrap(~wmo) + theme_bw()

pv_total

ggplot(vertical_profiles, aes(x = juld, y = depth, colour = total_pred)) + geom_point() + #scale_color_viridis(option = 'H') +
  scale_y_reverse(limits = c(500,0)) +
  scale_colour_viridis(option = 'A', limits = c(20,100), oob = scales::squish, na.value = 'grey', direction = -1) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'Predicted Ftotal') + facet_wrap(~wmo) + theme_bw()

```{r, flux distribution}

depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
flux_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, total_pred) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
  mutate(month = month.abb[month(juld)]) |> ungroup() #|> filter(month != 'May') # I filtered May because I had some troubles.. I need to fill some cases unseen by the float (eg. when they start profiling until 1000 m then go to 2000 -> ça fout un peu la merde..)

tmp <- flux_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_flux = mean(total_pred)) |> 
  mutate(cut_flux = cut(mean_flux, breaks = c(0,10,20,30,40,50,75,100,+Inf))) |> ungroup()

create_data_for_heatmap_bis <- function(float, df){
  # df <- tmp |> filter(wmo == float)
  # y <- unique(df$depth_bin)
  # if(float == 1902578){
  #   y <- y[1:15]
  # }
  # x <- unique(df$month)
  # data <- expand.grid(X=x, Y=y)
  # data$Z <- df$cut_flux
  # data$wmo <- float
  # data$month <- df$month
  df <- df |> filter(wmo == float) |> select(cut_flux, depth_bin, month) 
  y <- unique(df$depth_bin)
  df <- df|> pivot_wider(names_from = depth_bin, values_from = cut_flux) 
  x <- unique(df$month)
  data <- expand.grid(X=x, Y=y)
  data$Z <- as.vector(as.matrix(df |> select(-month)))
  data$wmo <- float
  return(data)
}

d1 <- create_data_for_heatmap_bis(6904241, tmp)
d2 <- create_data_for_heatmap_bis(1902578, tmp)
d3 <- create_data_for_heatmap_bis(4903634, tmp)
d4 <- create_data_for_heatmap_bis(6904240, tmp)

dd <- rbind(d1, d2, d3, d4) |> mutate(X = factor(X, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))) |>
  mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T))) |>
  mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578))) |>
  mutate(Z = factor(Z, levels = levels(tmp$cut_flux)))

ggplot(dd) + 
  geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Flux range', x = 'Month', y = 'Depth bin') +
  facet_wrap(~wmo) + scale_fill_viridis_d(option = 'A')

#### same plot mais pour les classes de tailles

```{r, TEST }

depth_bin <- c(seq(0,500,50), seq(600, 1000, 100), seq(1250, 2000, 250))
part_distrib <- vertical_profiles |> filter(depth < 2000) |> select(juld, depth, wmo, NP_Size_512) |> mutate(depth_bin = cut(depth, breaks = depth_bin)) |>
  mutate(month = month.abb[month(juld)]) |> ungroup() #|> filter(month != 'May') # I filtered May because I had some troubles.. I need to fill some cases unseen by the float (eg. when they start profiling until 1000 m then go to 2000 -> ça fout un peu la merde..)

tmp <- part_distrib |> group_by(wmo, depth_bin, month) |> summarize(mean_part = log10(mean(NP_Size_512))) |> 
  mutate(cut_part = cut(mean_part, breaks = c(seq(0,100,20),+Inf))) |> ungroup()

create_data_for_heatmap_bis <- function(float, df){
  df <- df |> filter(wmo == float) |> select(mean_part, depth_bin, month) 
  y <- unique(df$depth_bin)
  df <- df|> pivot_wider(names_from = depth_bin, values_from = mean_part) 
  x <- unique(df$month)
  data <- expand.grid(X=x, Y=y)
  data$Z <- as.vector(as.matrix(df |> select(-month)))
  data$wmo <- float
  return(data)
}

d1 <- create_data_for_heatmap_bis(6904241, tmp)
d2 <- create_data_for_heatmap_bis(1902578, tmp)
d3 <- create_data_for_heatmap_bis(4903634, tmp)
d4 <- create_data_for_heatmap_bis(6904240, tmp)

dd <- rbind(d1, d2, d3, d4) |> mutate(X = factor(X, levels = c('May','Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))) |>
  mutate(Y = factor(Y, levels = sort(unique(Y), decreasing = T))) |>
  mutate(wmo = factor(wmo, levels = c(6904240,4903634,6904241,1902578))) #|>
  #mutate(Z = factor(Z, levels = levels(tmp$cut_part)))

ggplot(dd) + 
  geom_tile(aes(X, Y, fill = Z), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Flux range', x = 'Month', y = 'Depth bin') +
  facet_wrap(~wmo) + scale_fill_viridis(option = 'A')

#### Compute fluxes with Guidi et al. 2008

A = 12.5, b = 3.81 fitted on 250 microns - 1.5 mm.

```{r, guidi 2008}

mid_DSE <- c(0.2892699,0.3644572,0.4591873,0.5785398,0.7289145,0.9183747,1.1570796,1.4578289)

guidi_spectra <- vertical_profiles |> select(NP_Size_203:NP_Size_1290)

# Guidi 2008
A <- 12.5
b <- 3.81
exp <- A * mid_DSE ** b
estimated_flux <- rowSums(t(t(guidi_spectra)*exp))
#estimated_flux

vertical_profiles$guidi_flux <- estimated_flux

ggplot(vertical_profiles, aes(x = juld, y = depth, colour = total_pred)) + geom_point() + #scale_color_viridis(option = 'H') +
  scale_y_reverse(limits = c(2000,0)) +
  scale_colour_viridis(option = 'A', limits = c(quantile(vertical_profiles$guidi_flux, probs = 0.25), quantile(vertical_profiles$guidi_flux, probs = 0.50)), oob = scales::squish, na.value = 'grey', direction = -1) +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'Predicted Guidi Flux') + facet_wrap(~wmo) + theme_bw()
```

#### Guidi vs. No Guidi Ha Diguedi

```{r, anomaly guidi}

ggplot(vertical_profiles |> mutate(ratio_flux = guidi_flux/total_pred)) + geom_point(aes(x = juld, y = depth, colour = ratio_flux)) +
  scale_y_reverse(limits = c(2000,0)) +
  scale_colour_viridis(option = 'A', limits = c(1,5), oob = scales::squish, na.value = 'grey') +
  scale_x_date(labels = date_format("%m"), date_breaks = '1 month') +
  labs(x = 'Time', y = 'Ratio Guidi/Total Model') + facet_wrap(~wmo) + theme_bw()

@@

@@
